Responses to Review Comments

### 2025-06-23 writeUps/paper.tex
more responses

\subsubsection{Paremetric Mixture Modeling}
%The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
Say an analyst assumes that, within principal strata, $Y_C$ is drawn from a parametric model with parameter vector $\bm{\delta}^s$, i.e., $f_{Y_C}(y|\st=s)=g^s(y)$. For instance, \citep{imbens1997bayesian} assumes $Y_C$ is normally distributed within strata, so $g^s(y)=\phi\left(\frac{y-\mucs}{\sigma_C^s}\right)$, where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations. 
Then the probability density of $Y_C$ may be written as
\begin{equation}\label{eq:mixtureModel}
 % f_{Y_C}(y)=Pr(\st=0)\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+Pr(\st=1)\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
 f_{Y_C}(y)=Pr(\st=0)g^0(y)+Pr(\st=1)g^1(y).
\end{equation}
The probabilities $Pr(\st=0)$ and $Pr(\st=1)$ may be estimated first using data from the treatment group, since Assumption \ref{ass:rand} implies that $Pr(\st=1|Z=1)=Pr(\st=1)$.
Given these estimates and \eqref{eq:mixtureModel}, parameters $\muc0$ and $\muc1$ may be estimated using maximum likelihood, method of moments, or Bayesian techniques.

@@

Like \geepers, \pmm is insensitive to interactions between $X$ and $S$. However, interactions between $X$ and $Z$ in the data generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.

@@

The simulation study suggests that none of the three methods evaluated, \geepers, \pmm, or \psw, dominates the others, and offers some guidance on when each estimator may be preferable. \psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability. When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option. When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$.

>>>>

\subsubsection{Parametric Mixture Modeling}
%The normal mixture modeling approach to estimating $\muc0$ and $\muc1$ \citep{imbens1997bayesian} assumes that, conditional on $\st$, $Y_C$ is normally distributed.
Say an analyst assumes that, within principal strata, $Y_C$ is drawn from a parametric model with parameter vector $\bm{\delta}^s$, i.e., $f_{Y_C}(y|\st=s)=g^s(y)$. For instance, \citep{imbens1997bayesian} assumes $Y_C$ is normally distributed within strata, so $g^s(y)=\phi\left(\frac{y-\mucs}{\sigma_C^s}\right)$, where $\phi(\cdot)$ is the standard normal density function, and $\sigma_C^0>0$ and $\sigma_C^1>0$ are standard deviations. 
Then the probability density of $Y_C$ may be written as
\begin{equation}\label{eq:mixtureModel}
 % f_{Y_C}(y)=Pr(\st=0)\phi\left(\frac{y-\muc0}{\sigma_C^0}\right)+Pr(\st=1)\phi\left(\frac{y-\muc1}{\sigma_C^1}\right)
 f_{Y_C}(y)=Pr(\st=0)g^0(y)+Pr(\st=1)g^1(y).
\end{equation}
The probabilities $Pr(\st=0)$ and $Pr(\st=1)$ may be estimated first using data from the treatment group, since Assumption \ref{ass:rand} implies that $Pr(\st=1|Z=1)=Pr(\st=1)$.
Given these estimates and \eqref{eq:mixtureModel}, parameters $\muc0$ and $\muc1$ may be estimated using maximum likelihood, method of moments, or Bayesian techniques.

@@

Like \geepers, \pmm is insensitive to interactions between $X$ and $S$. However, interactions between $X$ and $Z$ in the data-generating model skew the \pmm sampling distribution and reduce the coverage of its 95\% confidence intervals, even when $\alpha=0.5$.
When the residual distribution was uniformly distributed, the \pmm sampling distribution appeared bimodal, and its credible intervals exhibited severe undercoverage.

@@

\label{evalMetric}The simulation study suggests that none of the three methods evaluated---\geepers, \pmm, or \psw---dominates the others, and offers some guidance on when each estimator may be preferable. \psw is the most precise and requires no assumptions about the distribution of outcomes, but can be quite sensitive to violations of principal ignorability. When the determinants of $\st$ are well-understood and measured, \psw seems to be the best option. When principal ignorability does not hold, potential outcomes are conditionally normal, and there are no $X:Z$ interactions within principal strata, the \pmm estimator outperforms \geepers, though the difference decreases with $n$ or $\alpha$. \geepers may be the best choice when conditional normality may not hold, or if there are $X:Z$ interactions and $\bm{x}$ is sufficiently predictive of $\st$ (AUC $\ge$0.675, corresponding to $\alpha \ge 0.5$, may serve as a threshold). 
These suggestions may depend heavily on the simulation setup here; applying them in general scenarios requires more research.

### 2025-06-23 writeUps/responses.tex
more responses

\item \begin{itquote}    Would it be possible to offer some evaluation metric for GEEPERs, which would indicate, under some conditions researchers could possibly opt or seek other estimators?
\end{itquote}
This is a great suggestion. 
\item \begin{itquote}     Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental designs) could be partitioned into associative and dissociative effects with GEEPERs?
\end{itquote}
\item \begin{itquote}    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also make sure you have comments in the code so someone reading the article can match the narrative with the code.
\end{itquote}
\end{enumerate}

@@

Comments to the Author

This paper develops a new estimator for principal stratification based on M-estimation and stacked equations. The proposed method is designed to be more robust and simpler to use than standard methods. The method is, however, tailored to a very specific set up: one-way noncompliance. In general, principal stratification is a much more general framework that can be used for a wide variety of settings where there is conditioning on some post-treatment quantity.

The method is outlined, and then evaluated through a series of simulations, and then an application is provided. My primary criticism is that the new method doesn't appear to outperform extant methods to a compelling degree.  I take up this point in more detail first for the simulations and second for the empirical application.

Simulations.

In my view, the simulations should show the following:

--When the assumptions of extant methods are met, GEEPERS should do no worse.
--When the assumptions of extend methods are not met, GEEPERS should do better.

Given this criteria, it seems to me that the simulation should focus on the last three factors in the design.  That is, I would hold the sample size and alpha fixed.  Currently too many factors are varied at the same time to give a clear picture.  I would alter Table 1, to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of relaxing them simultaneously.  As it currently stands, it is quite hard to understand the relative performance of GEEPERs compared to extant methods. Careful relaxation of each assumption allows the reader to see how much bias results from a single factor.

@@

I thought there are a number of less than desirable parts to the application. First, the focus on bottoming out is less than ideal. This strikes me as a fairly rare phenomenon. I didn't find the concept very intuitive and had to read just the descriptive section several times to understand why one might be interested in a principal stratification analysis here.  This makes the methods seem not particularly general.  Why not focus on an application with straightforward oneway noncompliance? This is an area where principal stratification is often applied, and the concepts are well understood.

In addition, it isn't ideal that the ATE in each condition is not estimated very precisely, that is the CI appears to cover zero in all cases. Next, all the principal stratification estimates are all pretty close to the same value as the ATE. That serves as another reason this application isn't compelling. It is hard to care about an application where principal stratification doesn't provide much in the way of insights relative to the ATE. Next, the performance of GEEPERs seems to be quite similar to extant methods in each scenario. Overall, to my thinking, there is little to recommend the current application. Again, there is little here that demonstrates the value of GEEPERs.

To my mind, these problems could both be fixed.  That is, the simulations could be redesigned to provide more compelling evidence. Next, the authors may be able to identify are better application.  Therefore, the manuscript should be revised and resubmitted.

\end{document}

>>>>

\item \begin{itquote}    Would it be possible to offer some evaluation metric for GEEPERs, which would indicate, under some conditions researchers could possibly opt or seek other estimators?
\end{itquote}
This is a great suggestion. We added the paragraph beginning ``The simulation study suggests\dots'' on page \pageref{evalMetric} to give some guidance along those lines. 
\item \begin{itquote}     Quickly checked Frangakis and Rubin (2002). Do you think the effect of bottom-outers (or similar experimental designs) could be partitioned into associative and dissociative effects with GEEPERs?
\end{itquote}
This is correct, but only in a narrow technical sense. 
Under our notation, Frangakis and Rubin (2002) define dissociative effects as contrasts between $\yti$ and $Y_{Ci}$ for the subjects $i$ with $S_{Ci}=\sti$; associative effects are contrasts for subjects with $S_{Ci}\ne\sti$. A contrast between associative and dissociative effects ``measures the degree to which a causal effect of treatment on outcome occurs together with a causal effect of treatment on the surrogate."  Since our paper deals only with the one-way noncompliance case, we set $S_{Ci}=0$ for all $i$ (Assumption 1). In Section \ref{sec:fh2t}, we define $\st=0$ for non-bottom-outers and $\st=1$ for bottom-outers. Therefore, as you wrote, an effect for non-bottom-outers would be dissociative, while an effect for bottom-outers would be associative. However, this is, at least in part, an artifact of our notation. It's true that anyone assigned to control would, by necessity, request fewer than 11 bottom-out hints (the median in the treatment group)---$S_C=0$---and that treatment assignment affects bottom-out-hint usage less for non-bottom-outers than for bottom-outers. However, many non-bottom-outers assigned to treatment will request \emph{some} bottom-out hints, whereas students assigned to control request none---our choice to dichotomize hint requests masks variation that does not exist in the control group. Furthermore, our analysis would have been identical had we set $S_{Ci}$ equal to any other constant or set it to NA.

\item \begin{itquote}    It would be great if you could share the code for estimation and simulation in the next round of review (if there is). Also make sure you have comments in the code so someone reading the article can match the narrative with the code.
\end{itquote}
Full replication code is available at \url{FILL IN} 
\end{enumerate}

@@

\subsection{Simulations}
\begin{itquote}
    In my view, the simulations should show the following:
\begin{itemize}
   \item When the assumptions of extant methods are met, GEEPERS should do no worse.
   \item When the assumptions of extent methods are not met, GEEPERS should do better.
\end{itemize}
\end{itquote}
Depending on the meaning of ``do no worse,'' \geepers may or may not meet those criteria. 
In the scenarios we tested, \geepers is always less biased and its nominal confidence intervals achieve higher empirical coverage than the competitors we tested.
However, the competing methods often boast smaller standard errors than \geepers. 
This set of results is typical for semi- or non-parametric methods, which tend to be less efficient than their parametric counterparts when the latter's assumptions are met, but are, say, less biased when those assumptions are violated.

Unfortunately, \geepers may not even meet that criterion relative to \psw. Our simulation did not include scenarios in which principal ignorability held, but we expect that in such scenarios---depending on other factors---\psw may outperform \geepers in every respect. 
That said, principal ignorability is untestable and, in our opinion, rather strong. 
As we stated on page \pageref{evalMetric} (in the paragraph beginning ``The simulation study suggests''), researchers will have to decide in each application which assumptions are the most plausible, and choose estimators based on that decision.

\begin{itquote} Given this criterion, it seems to me that the simulation should focus on the last three factors in the design.  That is, I would hold the sample size and alpha fixed.  Currently, too many factors are varied at the same time to give a clear picture.  
I would alter Table 1 to indicate which factors yield assumption violations and then violate these assumptions in isolation instead of relaxing them simultaneously.  As it currently stands, it is quite hard to understand the relative performance of GEEPERs compared to extant methods. Careful relaxation of each assumption allows the reader to see how much bias results from a single factor. 
\end{itquote}
The results in Figure \ref{fig:boxplots} and in Table \ref{tab:coverage} now hold $n$ fixed at $n=500$ per group and completely cross the other factors, i.e. $\alpha=0.2$ or $0.5$, interactions between $\bm{x}$ and $\st$ or $Z$, and the distribution of the residuals.
Now it is apparent from Figure \ref{fig:boxplots} that, for instance, none of the methods are sensitive to interactions between $\bm{x}$ and $\st$, and that \pmm, but not the others, is sensitive to the distribution of the residuals.

In Table \ref{tab:coverage}, results when one or more assumptions have been violated are highlighted in red.

@@

I thought there were a number of less-than-desirable parts to the application. First, the focus on bottoming out is less than ideal. This strikes me as a fairly rare phenomenon. I didn't find the concept very intuitive and had to read just the descriptive section several times to understand why one might be interested in a principal stratification analysis here.  This makes the methods seem not particularly general.  Why not focus on an application with straightforward one-way noncompliance? This is an area where principal stratification is often applied, and the concepts are well understood.

In addition, it isn't ideal that the ATE in each condition is not estimated very precisely; that is, the CI appears to cover zero in all cases. Next, all the principal stratification estimates are pretty close to the same value as the ATE. That serves as another reason this application isn't compelling. It is hard to care about an application where principal stratification doesn't provide much in the way of insights relative to the ATE. Next, the performance of GEEPERs seems to be quite similar to extant methods in each scenario. Overall, to my thinking, there is little to recommend the current application. Again, there is little here that demonstrates the value of GEEPERs.

To my mind, these problems could both be fixed.  That is, the simulations could be redesigned to provide more compelling evidence. Next, the authors may be able to identify better applications.  Therefore, the manuscript should be revised and resubmitted.

\end{document}}
